Learning action models with minimal observability

### 2018-07-23 doc/aij18/learning_task.tex

Our approach to learning action models is to compile this learning task into a classical planning task with conditional effects.  Conditional effects allow us to compactly define actions whose effects depend on the current state. Supporting conditional effects is now a requirement of the IPC~\cite{vallati:IPC:AIM2015} and many classical planners cope with conditional effects without compiling them away.

@@

In section \ref{sec:motivation}, we exposed our reasons to solve the learning task $\Lambda=\tup{\mathcal{M},\tau}$ via compiling it into a classical planning task $P_{\Lambda}$. Our compilation scheme  builds upon the approach presented in \cite{aineto2018learning} but \FAMA comes up with a more general and flexible scheme able to capture any type of input plan trace.

The intuition behind the \FAMA compilation is that a solution plan $\pi_\Lambda$ to $P_{\Lambda}$ induces the output set of action models $\mathcal{M}'$ that solves $\Lambda$. Specifically, a solution plan $\pi_\Lambda$ serves two purposes:

@@

Finally, we will detail the composition of a solution plan $\pi_\Lambda$ to a planning task $P_\Lambda$ and the mechanism to extract the output set of action models $\mathcal{M}'$ from $\pi_\Lambda$. The plan of Figure~\ref{fig:plan-lplan} shows a solution to the task $P_{\Lambda}$ that encodes a learning task $\Lambda=\tup{\mathcal{M},\tau}$ for obtaining the action models of the {\em blocksworld} domain, where the models for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are already specified in $\mathcal{M}$. Therefore, this plan programs and validates the action model for {\tt\small stack}, using the input plan trace of Figure~\ref{fig:example-plans}. Plan steps $00-01$ program the preconditions of the {\tt\small stack} model, steps $02-06$ program the action model effects and steps $07-11$ is the plan postfix that validates the programmed model following the trace of Figure~\ref{fig:example-plans}.

>>>>

Conditional effects allow us to compactly define actions whose effects depend on the current state. Supporting conditional effects is now a requirement of the IPC~\cite{vallati:IPC:AIM2015} and many classical planners cope with conditional effects without compiling them away.

@@

In section \ref{sec:motivation}, we exposed our reasons to solve the learning task via compiling it into a classical planning task. Our compilation scheme  builds upon the approach presented in \cite{aineto2018learning} but \FAMA comes up with a more general and flexible scheme able to capture any type of input plan trace.

Our approach to solving a learning task $\Lambda=\tup{\mathcal{M},\tau}$ is to compile $\Lambda$ into a classical planning task $P_{\Lambda}$ with conditional effects. The intuition behind the \FAMA compilation is that a solution plan $\pi_\Lambda$ to $P_{\Lambda}$ induces the output set of action models $\mathcal{M}'$ that solves $\Lambda$. Specifically, a solution plan $\pi_\Lambda$ serves two purposes:

@@

Finally, we will detail the composition of a solution plan $\pi_\Lambda$ to a planning task $P_\Lambda$ and the mechanism to extract the output set of action models $\mathcal{M}'$ from $\pi_\Lambda$. The plan of Figure~\ref{fig:plan-lplan} shows a solution to the task $P_{\Lambda}$ that encodes a learning task $\Lambda=\tup{\mathcal{M},\tau}$ for obtaining the action models of the {\em blocksworld} domain, where the models for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are already specified in $\mathcal{M}$. Therefore, this plan programs and validates the action model for {\tt\small stack}, using the input plan trace of Figure~\ref{fig:example-plans}. Plan steps $00-01$ program the preconditions of the {\tt\small stack} model, steps $02-06$ program the action model effects and steps $07-11$ is the plan postfix that validates the programmed model following the plan trace of Figure~\ref{fig:example-plans}.

### 2018-07-23 doc/aij18/motivation.tex

\FAMA represents one step ahead towards learning action models without assuming observed actions. The main novelty of \FAMA with respect to other approaches lies in that our system is capable of handling \PO and \NO action sequences, which combined with \PO and \NO state trajectories, make the learning task more challenging (the actual length of the input plan trace becomes a priori unbound). This essentially brings one key difference: the transition between two given observed states may now involve more than one action; i.e., $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, with $k \geq 1$, $k$ unknown and unbound, and so the horizon of the input plan traces is no longer known now. Table \ref{tab:complex} shows when the worst case complexity of learning \strips\ action models becomes PSPACE-complete.

@@

\vspace{0.1cm}
Regarding the evaluation of the learned action models, we can observe in Table \ref{table:models_comparison2} that most of the approaches use a similar syntax-based metric that consists in (1) counting the missing and extra fluents that appear in the learned model wrt the GTM and (2), normalizing this error by the the total number of all the possible preconditions and effects of an action model. This is an \emph{optimistic} metric since error rates are not normalized by the size of the actual GTM. The set of preconditions and effects of the GTM is usually smaller than the set of all possible preconditions and effects hence, it turns out that these syntax-based metrics may output error rates below 100\% for totally wrong learned models. To overcome this limitation we propose to use two standard metrics from ML, {\em precision} and {\em recall}, that are frequently used in pattern recognition, information retrieval and binary classification~\cite{davis2006relationship}. %{\em Precision} and {\em recall} are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model.

Pure syntax-based evaluation metrics, like the ones mentioned in the above paragraph, can report low scores for learned models that are actually {\em sound} and {\em complete} but syntactically different from the GTM. Semantic evaluation metrics add a distinctive value over the syntactic ones, which is that they evaluate the learned model with a set of observations of plan executions and hence they are appropriate for scenarios where the GTM is not available. In this sense, \FAMA contributes also with a novel semantic-based error measure that builds upon the {\em precision} and {\em recall} metrics. Unlike the semantic metric used in \ARMS~\cite{yang2007learning}, our semantic version of {\em precision} and {\em recall} is not sensitive to the repetition of one same flaw in the model evaluation.

>>>>

\FAMA represents one step ahead towards learning action models without assuming observed actions. The main novelty of \FAMA with respect to other approaches lies in that our system is capable of handling \PO and \NO action sequences, which combined with \PO and \NO state trajectories, make the learning task more challenging. This essentially brings one key difference: the transition between two given observed states may now involve more than one action; i.e., $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, with $k \geq 1$, $k$ unknown and unbound, and so the horizon of the input plan traces is no longer known now. Table \ref{tab:complex} shows that the worst case complexity of learning \strips\ action models becomes PSPACE-complete when combining \PO and \NO state trajectories and action sequences.

@@

\vspace{0.1cm}
Regarding the evaluation of the learned action models, we can observe in Table \ref{table:models_comparison2} that most of the approaches use a similar syntax-based metric that consists in (1) counting the missing and extra fluents that appear in the learned model wrt the GTM and (2) normalizing this error by the the total number of all the possible preconditions and effects of an action model. This is an \emph{optimistic} metric since error rates are not normalized by the size of the actual GTM. The set of preconditions and effects of the GTM is usually smaller than the set of all possible preconditions and effects and thereby it turns out that these syntax-based metrics may output error rates below 100\% for totally wrong learned models. To overcome this limitation we propose to use two standard metrics from ML, {\em precision} and {\em recall}, that are frequently used in pattern recognition, information retrieval and binary classification~\cite{davis2006relationship}. %{\em Precision} and {\em recall} are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model.

Pure syntax-based evaluation metrics, like the ones mentioned in the above paragraph, can report low scores for learned models that are actually {\em sound} and {\em complete} but syntactically different from the GTM. Semantic evaluation metrics add a distinctive value over the syntactic ones, which is that they evaluate the learned model with a set of observations of plan executions and hence they are appropriate for scenarios where the GTM is not available. In this sense, \FAMA also contributes with a novel semantic-based error measure that builds upon the {\em precision} and {\em recall} metrics. Unlike the semantic metric used in \ARMS~\cite{yang2007learning}, our semantic version of {\em precision} and {\em recall} is not sensitive to the repetition of one same flaw in the model evaluation.
